Incentivizing Exploration with Unbiased Histories

### 2019-01-30 COLT 2019 Submission/llevel.tex
COLT submission
remove fullpage

\begin{theorem}
\label{thm:llevel}
The $L$-level recommendation policy gets regret $O\left(T^{2^{L-1}/(2^L-1)} \log^2(T) \right)$ for $L \leq \log(\ln(T)/\log(\NG^4))$. In particular, if we pick $L = \log(\ln(T)/\log(\NG^4))$, the regret is $O(T^{1/2} polylog(T))$. 
\end{theorem}

@@

We prove the above via induction on $l$. %For notation convenience, define $\bar{\mu}^{l,u}_a$  to be the empirical mean of arm $a$ pulls in the history observed by agents in groups $G_{l,1,u},...,G_{l,\NG,u}$ and $\Gamma_{l,1,u},...,\Gamma_{l,\NG,u}$ (they observe the same history).
We start by the base case when $l=2$. For each arm $a$, $W_4$ implies there exists $u_a$ such that $W^{u_a,a,high}_4$ and $W^{u_a,3-a,low}_4$ happen. For an agent $t$  in groups $G_{2,1,u_a},...,G_{2,\NG,u_a}$ and $\Gamma_{2,1,u_a},...,\Gamma_{2,\NG,u_a}$.
$W_4^{u_a,a,high}$,  $W_1^{a,u_a,v}$ and $W_2$ together imply that 
\begin{align*}
\bar{\mu}_a ^t &\geq \mu_a + \left(\fdpN T_1\NG \cdot \frac{1}{\sqrt{\fdpN T_1\NG}} - \GdT \sqrt{T_1\log(T)} \NG\cdot \sqrt{\frac{3\log(T)}{ \GdT \sqrt{T_1\log(T)}\NG}} \right) \cdot \frac{1}{(\fdpN T_1+ \GdT \sqrt{T_1\log(T)})\NG} \\
&> \mu_a + \frac{1}{4\sqrt{\fdpN T_1\NG}}.
\end{align*}
The second last inequality holds when $T$ is larger than some constant.
Similarly, we also have
\[
\bar{\mu}_{3-a}^t< \mu_{3-a}   - \frac{1}{4\sqrt{\fdpN[3-a] T_1\NG}}.
\]
Then we have
\begin{align*}
\bar{\mu}^t_a - \bar{\mu}^t_{3-a} &> \mu_a - \mu_{3-a} + \frac{1}{4\sqrt{\fdpN T_1\NG}} + \frac{1}{4\sqrt{\fdpN[3-a] T_1\NG}}\\
&\geq -\varepsilon_1+ \frac{1}{4\sqrt{\fdpN T_1\NG}} + \frac{1}{4\sqrt{\fdpN[3-a] T_1\NG}}\\
&\geq \frac{1}{8\sqrt{\fdpN T_1\NG}} + \frac{1}{8\sqrt{\fdpN[3-a] T_1\NG}}.
\end{align*}
By Assumption \ref{ass:embehave}, we have
\begin{align*}
\hat{\mu}_a^t - \hat{\mu}_{3-a}^t &> \bar{\mu}^t_a - \bar{\mu}^t_{3-a} -  \frac{\estC}{\sqrt{\fdpN T_1\NG/2}} - \frac{\estC}{\sqrt{\fdpN[3-a] T_1\NG/2}}\\
&> \frac{1}{8\sqrt{\fdpN T_1\NG}} + \frac{1}{8\sqrt{\fdpN[3-a] T_1\NG}} -   \frac{\estC}{\sqrt{\fdpN T_1\NG/2}} - \frac{\estC}{\sqrt{\fdpN[3-a] T_1\NG/2}}\\
&>0.
\end{align*}
The last inequality holds since $\estC$ is a small enough constant defined in Assumption \ref{ass:embehave}. Therefore we know agents in groups $G_{2,1,u_a},...,G_{2,\NG,u_a}$ and $\Gamma_{2,1,u_a},...,\Gamma_{2,\NG,u_a}$ all pull arm $a$.

@@

\begin{itemize}
\item $\varepsilon_{l-1} \NG \leq \mu_1 - \mu_2 \leq \varepsilon_{l-2}$ for $l = 2$(i.e. $\varepsilon_1\NG\leq \mu_1 - \mu_2 \leq \varepsilon_0$): We know agents in level at least 2 will observe at least $\fdpN T_1/2$ pulls of arm $a$ for $a \in \{1,2\}$. By $W_2$, for any agent in level at least 2, we have
\[
|\bar{\mu}_a^t - \mu_a| \leq \sqrt{\frac{3\log(T)}{\NG\fdpN T_1/2}}.
\]
By Assumption \ref{ass:embehave}, we have
\begin{align*}
\hat{\mu}_1^t - \hat{\mu}_2^t &\geq \bar{\mu}_1^t - \bar{\mu}_2^t - \frac{\estC}{\sqrt{\NG\fdpN[1]T_1/2}} -  \frac{\estC}{\sqrt{\NG\fdpN[2]T_1/2}}\\
&\geq \mu_1 -\mu_2-\sqrt{\frac{3\log(T)}{\NG\fdpN[1]T_1/2}} - \sqrt{\frac{3\log(T)}{\NG\fdpN[2]T_1/2}}- \frac{\estC}{\sqrt{\NG\fdpN[1]T_1/2}} -  \frac{\estC}{\sqrt{\NG\fdpN[2]T_1/2}}\\
&\geq\frac{\sqrt{\NG}}{4\sqrt{\fdpN[1]T_1}} +  \frac{\sqrt{\NG}}{4\sqrt{\fdpN[2]T_1}}-\sqrt{\frac{3\log(T)}{\NG\fdpN[1]T_1/2}} - \sqrt{\frac{3\log(T)}{\NG\fdpN[2]T_1/2}}- \frac{\estC}{\sqrt{\NG\fdpN[1]T_1/2}} -  \frac{\estC}{\sqrt{\NG\fdpN[2]T_1/2}}\\
&>0.
\end{align*}
Therefore agents in level at least 2 will all pull arm 1.

@@

\item $\varepsilon_{l-2} < \mu_1 - \mu_2 < \varepsilon_{l-2}\NG$ for $l =3$ (i.e. $\varepsilon_1 < \mu_1 - \mu_2 < \varepsilon_1\NG$): By Claim \ref{clm:l2_explore}, for any agent $t$ in level at least $3$, that agent will observe at least $T_1\fdpN \NG^2/2$ arm $a$ pulls (just from the first level). By $W_2$, we have
\[
|\bar{\mu}_a^t - \mu_a| \leq \sqrt{\frac{3\log(T)}{\NG^2\fdpN T_1/2}}.
\]
By Assumption \ref{ass:embehave}, we have
\begin{align*}
\hat{\mu}_1^t - \hat{\mu}_2^t &\geq \bar{\mu}_1^t - \bar{\mu}_2^t - \frac{\estC}{\sqrt{\NG^2\fdpN[1]T_1/2}}-\frac{\estC}{\sqrt{\NG^2\fdpN[2]T_1/2}}  \\
&\geq \mu_1 -\mu_2 -  \sqrt{\frac{3\log(T)}{\NG^2\fdpN[1]T_1/2}}- \sqrt{\frac{3\log(T)}{\NG^2\fdpN[2]T_1/2}}- \frac{\estC}{\sqrt{\NG^2\fdpN[1]T_1/2}}-\frac{\estC}{\sqrt{\NG^2\fdpN[2]T_1/2}}  \\
&\geq\frac{1}{4\sqrt{\NG\fdpN[1]T_1}} + \frac{1}{4\sqrt{\NG\fdpN[2]T_1}}  -  \sqrt{\frac{3\log(T)}{\NG^2\fdpN[1]T_1/2}}- \sqrt{\frac{3\log(T)}{\NG^2\fdpN[2]T_1/2}}- \frac{\estC}{\sqrt{\NG^2\fdpN[1]T_1/2}}-\frac{\estC}{\sqrt{\NG^2\fdpN[2]T_1/2}}  \\
&>0.
\end{align*}
Therefore agents in level at least 3 will all pull arm 1.

@@

Finally we proceed to bound the regret conditioned on the intersection of clean events happens. The proofs of Theorem \ref{thm:llevel} and Corollary \ref{cor:llevel} bound it by consider the regret from pulling the suboptimal arm (i.e. arm 2). When extending to more than 2 arms, we can do the exactly same argument for all arms except arm 1. This will blow up the regret by a factor of $(K-1)$ which is a constant.
\qed

>>>>

\begin{theorem}
\label{thm:llevel}
The $L$-level recommendation policy gets regret $O\left(T^{2^{L-1}/(2^L-1)} \log^2(T) \right)$ for \\ $L \leq \log(\ln(T)/\log(\NG^4))$. In particular, if we pick $L = \log(\ln(T)/\log(\NG^4))$, the regret is $O(T^{1/2} polylog(T))$. 
\end{theorem}

@@

We prove the above via induction on $l$. %For notation convenience, define $\bar{\mu}^{l,u}_a$  to be the empirical mean of arm $a$ pulls in the history observed by agents in groups $G_{l,1,u},...,G_{l,\NG,u}$ and $\Gamma_{l,1,u},...,\Gamma_{l,\NG,u}$ (they observe the same history).
We start by the base case when $l=2$. For each arm $a$, $W_4$ implies there exists $u_a$ such that $W^{u_a,a,high}_4$ and $W^{u_a,3-a,low}_4$ happen. For an agent $t$  in groups $G_{2,1,u_a},...,G_{2,\NG,u_a}$ and $\Gamma_{2,1,u_a},...,\Gamma_{2,\NG,u_a}$.
$W_4^{u_a,a,high}$,  $W_1^{a,u_a,v}$ and $W_2$ together imply that 
\begin{align*}
\bar{\mu}_a ^t &\geq \mu_a + \left(\fdpN T_1\NG \cdot \frac{1}{\sqrt{\fdpN T_1\NG}} - \GdT \sqrt{T_1\log(T)} \NG\cdot \sqrt{\frac{3\log(T)}{ \GdT \sqrt{T_1\log(T)}\NG}} \right) \\
&~~~~~~~~~~~\cdot \frac{1}{(\fdpN T_1+ \GdT \sqrt{T_1\log(T)})\NG} \\
&> \mu_a + \frac{1}{4\sqrt{\fdpN T_1\NG}}.
\end{align*}
The second last inequality holds when $T$ is larger than some constant.
Similarly, we also have
\[
\bar{\mu}_{3-a}^t< \mu_{3-a}   - \frac{1}{4\sqrt{\fdpN[3-a] T_1\NG}}.
\]
Then we have
\begin{align*}
\bar{\mu}^t_a - \bar{\mu}^t_{3-a} &> \mu_a - \mu_{3-a} + \frac{1}{4\sqrt{\fdpN T_1\NG}} + \frac{1}{4\sqrt{\fdpN[3-a] T_1\NG}}\\
&\geq -\varepsilon_1+ \frac{1}{4\sqrt{\fdpN T_1\NG}} + \frac{1}{4\sqrt{\fdpN[3-a] T_1\NG}}\\
&\geq \frac{1}{8\sqrt{\fdpN T_1\NG}} + \frac{1}{8\sqrt{\fdpN[3-a] T_1\NG}}.
\end{align*}
By Assumption \ref{ass:embehave}, we have
\begin{align*}
\hat{\mu}_a^t - \hat{\mu}_{3-a}^t &> \bar{\mu}^t_a - \bar{\mu}^t_{3-a} -  \frac{\estC}{\sqrt{\fdpN T_1\NG/2}} - \frac{\estC}{\sqrt{\fdpN[3-a] T_1\NG/2}}\\
&> \frac{1}{8\sqrt{\fdpN T_1\NG}} + \frac{1}{8\sqrt{\fdpN[3-a] T_1\NG}} -   \frac{\estC}{\sqrt{\fdpN T_1\NG/2}} - \frac{\estC}{\sqrt{\fdpN[3-a] T_1\NG/2}}\\
&>0.
\end{align*}
The last inequality holds since $\estC$ is a small enough constant defined in Assumption \ref{ass:embehave}. Therefore we know agents in groups $G_{2,1,u_a},...,G_{2,\NG,u_a}$ and $\Gamma_{2,1,u_a},...,\Gamma_{2,\NG,u_a}$ all pull arm $a$.

@@

\begin{itemize}
\item $\varepsilon_{l-1} \NG \leq \mu_1 - \mu_2 \leq \varepsilon_{l-2}$ for $l = 2$(i.e. $\varepsilon_1\NG\leq \mu_1 - \mu_2 \leq \varepsilon_0$): We know agents in level at least 2 will observe at least $\fdpN T_1/2$ pulls of arm $a$ for $a \in \{1,2\}$. By $W_2$, for any agent in level at least 2, we have
\[
|\bar{\mu}_a^t - \mu_a| \leq \sqrt{\frac{3\log(T)}{\NG\fdpN T_1/2}}.
\]
By Assumption \ref{ass:embehave}, we have
\begin{align*}
\hat{\mu}_1^t - \hat{\mu}_2^t &\geq \bar{\mu}_1^t - \bar{\mu}_2^t - \frac{\estC}{\sqrt{\NG\fdpN[1]T_1/2}} -  \frac{\estC}{\sqrt{\NG\fdpN[2]T_1/2}}\\
&\geq \mu_1 -\mu_2-\sqrt{\frac{3\log(T)}{\NG\fdpN[1]T_1/2}} - \sqrt{\frac{3\log(T)}{\NG\fdpN[2]T_1/2}}- \frac{\estC}{\sqrt{\NG\fdpN[1]T_1/2}} -  \frac{\estC}{\sqrt{\NG\fdpN[2]T_1/2}}\\
&\geq\frac{\sqrt{\NG}}{4\sqrt{\fdpN[1]T_1}} +  \frac{\sqrt{\NG}}{4\sqrt{\fdpN[2]T_1}}-\sqrt{\frac{3\log(T)}{\NG\fdpN[1]T_1/2}} - \sqrt{\frac{3\log(T)}{\NG\fdpN[2]T_1/2}}\\
&~~~~~~- \frac{\estC}{\sqrt{\NG\fdpN[1]T_1/2}} -  \frac{\estC}{\sqrt{\NG\fdpN[2]T_1/2}}\\
&>0.
\end{align*}
Therefore agents in level at least 2 will all pull arm 1.

@@

\item $\varepsilon_{l-2} < \mu_1 - \mu_2 < \varepsilon_{l-2}\NG$ for $l =3$ (i.e. $\varepsilon_1 < \mu_1 - \mu_2 < \varepsilon_1\NG$): By Claim \ref{clm:l2_explore}, for any agent $t$ in level at least $3$, that agent will observe at least $T_1\fdpN \NG^2/2$ arm $a$ pulls (just from the first level). By $W_2$, we have
\[
|\bar{\mu}_a^t - \mu_a| \leq \sqrt{\frac{3\log(T)}{\NG^2\fdpN T_1/2}}.
\]
By Assumption \ref{ass:embehave}, we have
\begin{align*}
&~~~\hat{\mu}_1^t - \hat{\mu}_2^t\\
 &\geq \bar{\mu}_1^t - \bar{\mu}_2^t - \frac{\estC}{\sqrt{\NG^2\fdpN[1]T_1/2}}-\frac{\estC}{\sqrt{\NG^2\fdpN[2]T_1/2}}  \\
&\geq \mu_1 -\mu_2 -  \sqrt{\frac{3\log(T)}{\NG^2\fdpN[1]T_1/2}}- \sqrt{\frac{3\log(T)}{\NG^2\fdpN[2]T_1/2}}- \frac{\estC}{\sqrt{\NG^2\fdpN[1]T_1/2}}-\frac{\estC}{\sqrt{\NG^2\fdpN[2]T_1/2}}  \\
&\geq\frac{1}{4\sqrt{\NG\fdpN[1]T_1}} + \frac{1}{4\sqrt{\NG\fdpN[2]T_1}}  -  \sqrt{\frac{3\log(T)}{\NG^2\fdpN[1]T_1/2}}- \sqrt{\frac{3\log(T)}{\NG^2\fdpN[2]T_1/2}}\\
&~~~~~~~~- \frac{\estC}{\sqrt{\NG^2\fdpN[1]T_1/2}}-\frac{\estC}{\sqrt{\NG^2\fdpN[2]T_1/2}}  \\
&>0.
\end{align*}
Therefore agents in level at least 3 will all pull arm 1.

@@

Finally we proceed to bound the regret conditioned on the intersection of clean events happens. The proofs of Theorem \ref{thm:llevel} and Corollary \ref{cor:llevel} bound it by consider the regret from pulling the suboptimal arm (i.e. arm 2). When extending to more than 2 arms, we can do the exactly same argument for all arms except arm 1. This will blow up the regret by a factor of $(K-1)$ which is a constant.
%\qed
